---
title: "Credit Analysis: Loan Default Prediction"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: scroll
    theme: cosmo
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(readr)
library(highcharter)
library(dplyr)
library(plotly)

setwd('C:/Users/Jimit/Desktop/FE550/FE550Proj')

library(highcharter)
load("CScore.rda")
load("post.RData")
load("during.RData")
load("pre.RData")
load("dfAfter.RData")


x <- rbind(d2000, d2001)$CSCORE_B

library(plotly)
D1=density(na.omit(rbind(d2000, d2001)$CSCORE_B),2)
D2=density(na.omit(rbind(d2002, d2003)$CSCORE_B),2)
D3=density(na.omit(rbind(d2004, d2005)$CSCORE_B),2)
D4=density(na.omit(rbind(d2006, d2007)$CSCORE_B),2)
D5=density(na.omit(rbind(d2008, d2009)$CSCORE_B),2)
D6=density(na.omit(rbind(d2010, d2011)$CSCORE_B),2)
D7=density(na.omit(rbind(d2012, d2013)$CSCORE_B),2)
D8=density(na.omit(rbind(d2014, d2015)$CSCORE_B),2)
D9=density(na.omit(rbind(d2016, d2017)$CSCORE_B),2)


df1=data.frame(cbind(x = D1$x, z = rep("2000-2001", 512), y = D1$y))
df2=data.frame(cbind(x = D2$x, z = rep("2002-2003", 512), y = D2$y))
df3=data.frame(cbind(x = D3$x, z = rep("2004-2005", 512), y = D3$y))
df4=data.frame(cbind(x = D4$x, z = rep("2006-2007", 512), y = D4$y))
df5=data.frame(cbind(x = D5$x, z = rep("2008-2009", 512), y = D5$y))
df6=data.frame(cbind(x = D6$x, z = rep("2010-2011", 512), y = D6$y))
df7=data.frame(cbind(x = D7$x, z = rep("2012-2013", 512), y = D7$y))
df8=data.frame(cbind(x = D8$x, z = rep("2014-2015", 512), y = D8$y))
df9=data.frame(cbind(x = D9$x, z = rep("2016-2017", 512), y = D9$y))



df=rbind(df1,df2,df3,df4,df5,df6,df7,df8,df9)
df=as.data.frame(df)
df$z=as.factor(df$z)

load("df2.RData")

```

Visualization {data-orientation=rows data-icon="fa-bar-chart"}
=====================================

## Row 1 {data-height=120}

### Number of Loans Pre Financial Crisis

```{r}
valueBox("16.3 Million", color="rgb(100,100,100)")
```

### Number of Loans During Financial Crisis

```{r}
valueBox("11.2 Million", color="rgb(200,100,100)")
```

### Number of Loans Post Financial Crisis

```{r}
valueBox("11.4 Million", color="#154360")
```    

## Row 2 {data-height=750}  

### Comparison of Borrower's Credit Scores
    
```{r}

x <- na.omit(df_pre_crisis$CSCORE_B)
y <- na.omit(df_dur_crisis$CSCORE_B)
z <- na.omit(df_post_crisis$CSCORE_B)

hchart(density(x, bw=7), area = F, name = "Pre Crisis") %>%
  hc_add_series(density(y, bw=7), area = F, name = "During Crisis") %>%
  hc_add_series(density(z, bw=7), area = F, name = "Post Crisis") %>% 
  hc_yAxis(title=list(text='Frequency')) %>% 
  hc_xAxis(title=list(text='Credit Score')) %>%
  hc_add_theme(hc_theme_smpl())

```

## Row 3 {data-height=750}  

### Comparison of Borrower's Interest Rate
    
```{r}

x <- na.omit(df_pre_crisis$ORIG_RT.x)
y <- na.omit(df_dur_crisis$ORIG_RT.x)
z <- na.omit(df_post_crisis$ORIG_RT.x)

hchart(density(x, bw=1), area = F, name = "Pre Crisis") %>%
  hc_add_series(density(y, bw=1), area = F, name = "During Crisis") %>%
  hc_add_series(density(z, bw=1), area = F, name = "Post Crisis") %>% 
  hc_yAxis(title=list(text='Frequency')) %>% 
  hc_xAxis(title=list(text='Interest Rate')) %>%
  hc_add_theme(hc_theme_smpl())

```

## Row 4 {data-height=750}  

### Comparison of Borrower's Debt to Income Ratio (DTI)
    
```{r}

x <- na.omit(df_pre_crisis$DTI)
y <- na.omit(df_dur_crisis$DTI)
z <- na.omit(df_post_crisis$DTI)

hchart(density(x, bw=2), area = F, name = "Pre Crisis") %>%
  hc_add_series(density(y, bw=2), area = F, name = "During Crisis") %>%
  hc_add_series(density(z, bw=2), area = F, name = "Post Crisis") %>% 
  hc_yAxis(title=list(text='Frequency')) %>% 
  hc_xAxis(title=list(text='DTI')) %>%
  hc_add_theme(hc_theme_smpl())

```

## Row 5 {data-height=750}  

### Comparison of Borrower's Original Loan to Value (OLTV)
    
```{r}

x <- na.omit(df_pre_crisis$OLTV)
y <- na.omit(df_dur_crisis$OLTV)
z <- na.omit(df_post_crisis$OLTV)

hchart(density(x, bw=2), area = F, name = "Pre Crisis") %>%
  hc_add_series(density(y, bw=2), area = F, name = "During Crisis") %>%
  hc_add_series(density(z, bw=2), area = F, name = "Post Crisis") %>% 
  hc_yAxis(title=list(text='Frequency')) %>% 
  hc_xAxis(title=list(text='OLTV')) %>%
  hc_add_theme(hc_theme_smpl())

```

## Row 6 {data-height=950}  

### Distribution of Credit Score
    
```{r}


axx <- list(
  title = "Credit Score"
)

axy <- list(
  title = "Year"
)

axz <- list(
  title = "Frequency"
)

plot_ly(df, x = ~x, y = ~z, z = ~y, group = ~z, type = "scatter3d", 
        mode = "lines", color=~z, line = list(width = 6))  %>%
  layout(title = "Distribution of Borrower's  Credit Score", scene = list(xaxis=axx,yaxis=axy,zaxis=axz))


```

## Row 7 {data-height=750} 

### Credit Score
    
```{r}
ggplot(databox, aes(x=year, y=CreditScore, fill=Default)) + 
  geom_boxplot() +
  xlab('Year') +
  ylab('Credit Score') +
  ggtitle('Boxplot: Credit Score') +
  theme_minimal()
```

### Interest Rate
    
```{r}
ggplot(databox, aes(x=year, y=OrInterestRate, fill=Default)) + 
  geom_boxplot() +
  xlab('Year') +
  ylab('Interest Rate') +
  ggtitle('Boxplot: Interest Rate') +
  theme_minimal()
```

## Row 8 {data-height=750} 

### Debt to Income
    
```{r}
ggplot(databox, aes(x=year, y=DTIRat, fill=Default)) + 
  geom_boxplot() +
  xlab('Year') +
  ylab('DTI') +
  ggtitle('Boxplot: Debt to Income') +
  theme_minimal()
```

### Loan to Value
    
```{r}
ggplot(databox, aes(x=year, y=OrLTV, fill=Default)) + 
  geom_boxplot() +
  xlab('Year') +
  ylab('LTV') +
  ggtitle('Boxplot: Loan to Value') +
  theme_minimal()
```

## Row 9 {data-height=750} 

### Purpose of Loan versus Loan Default
    
```{r}
x=df2$LoanPurpose[df2$Default==1]
purpose_default=summary(x)[-4]
purpose_default=purpose_default/sum(purpose_default)*100

y=df2$LoanPurpose[df2$Default==0]
purpose_notdefault=summary(y)[-4]
purpose_notdefault=purpose_notdefault/sum(purpose_notdefault)*100



Loan.Purpose <- c("Cash-out Refinance", "Purchase", "Refinance")

SF_Zoo <- purpose_default
LA_Zoo <- purpose_notdefault
data <- data.frame(Loan.Purpose, LA_Zoo, SF_Zoo)


p <- data %>% 
  plot_ly() %>%
  add_trace(x = ~Loan.Purpose, y = ~LA_Zoo, type = 'bar', name = 'Not Default',
            marker = list(color = 'rgba(219, 64, 82, 0.7)',
                          line = list(color = 'rgba(219, 64, 82, 1.0)', width = 1.5))) %>%
  add_trace(x = ~Loan.Purpose, y = ~SF_Zoo, type = 'bar', name = 'Default', 
            marker = list(color = 'rgba(55, 128, 191, 0.7)',
                          line = list(color = 'rgba(55, 128, 191, 1.0)', width = 1.5))) %>%
  layout(title = "Purpose of Loan versus Loan Default",
         barmode = 'group',
         xaxis = list(title = "Purpose of Loan"),
         yaxis = list(title = "Percentage"))
p
```

### Occupancy Status of Property versus Loan Default
    
```{r}
x=df2$OccStatus[df2$Default==1]
OccStatus_default=summary(x)
OccStatus_default=OccStatus_default/sum(OccStatus_default)*100

y=df2$OccStatus[df2$Default==0]
OccStatus_notdefault=summary(y)
OccStatus_notdefault=OccStatus_notdefault/sum(OccStatus_notdefault)*100

Occupation.Status <- c("Investment", "Principal Residence", "Second Home")
LA_Zoo <- OccStatus_notdefault
SF_Zoo <- OccStatus_default
data <- data.frame(Occupation.Status, LA_Zoo, SF_Zoo)

p <- data %>% 
  plot_ly() %>%
  add_trace(x = ~Occupation.Status, y = ~LA_Zoo, type = 'bar', name = 'Not Default',
            marker = list(color = 'rgba(219, 64, 82, 0.7)',
                          line = list(color = 'rgba(219, 64, 82, 1.0)', width = 1.5))) %>%
  add_trace(x = ~Occupation.Status, y = ~SF_Zoo, type = 'bar', name = 'Default', 
            marker = list(color = 'rgba(55, 128, 191, 0.7)',
                          line = list(color = 'rgba(55, 128, 191, 1.0)', width = 1.5))) %>%
  layout(title = "Occupancy Status of Property versus Loan Default",
         barmode = 'group',
         xaxis = list(title = "Occupancy Status"),
         yaxis = list(title = "Percentage"))
p

```


## Row 10 {data-height=800}  

### Property Type versus Loan Default
    
```{r}

x=df2$PropertyType[df2$Default==1]
Property_default=summary(x)
Property_default=Property_default/sum(Property_default)*100

y=df2$PropertyType[df2$Default==0]
Property_notdefault=summary(y)
Property_notdefault=Property_notdefault/sum(Property_notdefault)*100

Property.Type <- c("Condo", "Cooperative Share", "Manufactured Housing", "Planned Urban Development", "Single Family")
LA_Zoo <- Property_notdefault
SF_Zoo <- Property_default
data <- data.frame(Property.Type, LA_Zoo, SF_Zoo)



p <- data %>% 
  plot_ly() %>%
  add_trace(x = ~Property.Type, y = ~LA_Zoo, type = 'bar', name = 'Not Default',
            marker = list(color = 'rgba(219, 64, 82, 0.7)',
                          line = list(color = 'rgba(219, 64, 82, 1.0)', width = 1.5))) %>%
  add_trace(x = ~Property.Type, y = ~SF_Zoo, type = 'bar', name = 'Default', 
            marker = list(color = 'rgba(55, 128, 191, 0.7)',
                          line = list(color = 'rgba(55, 128, 191, 1.0)', width = 1.5))) %>%
  layout(title = "Property Type versus Loan Default",
         barmode = 'group',
         xaxis = list(title = "Property Type"),
         yaxis = list(title = "Percentage"))
p
```



